---
title: "Assignment_2"
Author: "Brianna Viñas" 
Date: 10/04/2021
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
library(dplyr)
library(caret)
library(ISLR)
library(FNN)
library(gmodels)
library(dummies)

#I am going to read the .csv file in R. I will name the variable UniBank. 
```{r}
UniBank<- read.csv("UniversalBank.csv")

#Now, I will proceed to remove the ID and zip code columns. 

UBR<- UniBank [, c(-1,-5)]
str(UBR)
```

#Next, I will create the dummy variable for the Education Column. 
```{r} 
Educate<- dummy(UBR$Education)
tmp<- cbind(UBR, Educate)
head(tmp)
```
#For this step, I will now take away the Education Column. 
```{r}
UBank<- tmp[c(-6)]
head(UBank)
```

#This is where I will partition the data. Training =60% and the Validation= 40%.
```{r}
set.seed(18)
Index <- createDataPartition(UBank$Income, p=0.6, list = FALSE)

training_data <-UBank[Index,]
dim(training_data)

validation_data<-UBank[Index,]
dim(validation_data)
```
#The data of the .csv will now be stabsardized using normalization. 
```{r}
norm_model<-preProcess(training_data, method= c("center", "scale"))
head(norm_model)
train_nf <- predict(norm_model, training_data)
validation_nf<-predict(norm_model, validation_data)
total_nf<- predict(norm_model,UBank)

summary(total_nf)
summary(train_nf)
summary(validation_nf)

set.seed(15)

training_data<-train_nf[-7]
head(training_data)
train_outcome<-factor(training_data[,7], levels= c (0,1), labels= C ("reject", "accept"))
head(train_outcome)
valid_data<-validation_nf[-7]
head(valid_data)
validation_outcome<-factor(validation_data[7], levels = C (0,1), labels = C("reject","accept"))
head(validation_outcome)
total_data<-total_nf[-7]
total_outcome<-factor(UBank[,7], levels = c(0,1), labels = c("reject", "accept"))
head(total_outcome)
```
# Question 1 Age= 40, Experience=10, Income=84,Family= 2, CCAvg=2, Education_1= 0, Education_2=1, Education_3=0, Mortgage=0,Securties Account=0, CD Account = 0,Online=1, Credit Card= 1. Perform a K-NN Classification with all predicators except ID and zipcode using k=1.Remember to transform categorical predicators with more than two categories into dummy variables first. Specify the success class as 1 (loan acceptance), and use the default cutoff vaule of 0.5. How would this customer be classified? 

#Answer Question 1: 
```{r}
TestTraining_Data <-c(40,10,84,2,2,0,0,0,1,1,0,1,0)
knn_test<- knn(train_data, TestTraining_Data,cl=train_outcome, k=1, prob= TRUE)
knn_test  

```
#Thus, the customer above would be classified as 0 as its nearest neighbors are classified as 0. 

# Question 2 What is a choice of k that balances overfitting and ignoring the predictor information? 

#Answer to Question 2: 
```{r}
bestk<-data.frame(k= seq(1,55,1), accuracy = rep(0,55)) 
head(bestk)
for (i in 1:55){
  knn.pred<- knn(train_data,valid_data, cl= train_outcome, k=i)
  bestk[i,2] <- confusionMatrix(knn.pred,validation_outcome)$overall [1]
}
head(bestk)
bestk_fit<-bestk[which.max(bestk$accuracy),]
bestk_fit
```
#Here the optimal value of k=4. 

#Question # 3 Show the confusion matrix for the validation data that results from using the best k. 

#Answer to Question 3
```{r}
knn.pred<-knn(train_data,valid_data, cl= train_outcome, k=bestk_fit$k, prob= TRUE)
CrossTable(validation_outcome,knn.pred)
```
#Question 4 Consider the following customer: Age= 40, Experience=10, Income=84,Family= 2, CCAvg=2, Education_1= 0, Education_2=1, Education_3=0, Mortgage=0,Securties Account=0, CD Account = 0,Online=1, Credit Card= 1. Classify the customer using the best k. 

#Answer to Question 4 
```{r}
TestTraining_Data<- C(40,10,84,2,2,0,0,0,1,1,0,1,0)
Bestfitknn <-knn (train_data, TestTraining_Data, cl=train_outcome, k=bestk_fit$k,prob= TRUE)
(Bestfitknn)

totalknn<- knn(train_data,total_data,cl=train_outcome,k=bestk_fit$k,prob =TRUE)
CrossTable(total_outcome, totalknn)
```


#Question 5 Repartition the data, this time training, validation, and test sets (50%), (30%), (20%). Apply the K-NN method with the k chosen above. Compare the confusion matrix of the test set with that of the training and validation sets. Comment on their differences and the reason. 

#Answer to Question 5
```{r}
IndexNew1 <- createDataPartition(UBank$Income, p=0.5, list= FALSE)
training_data2 = UBank [IndexNew1,]

Remdata<- UBank[-IndexNew1,]
IndexNew2 <- createDataPartition(Remdata$Income, p=0.6, list= FALSE) 
validation_data2 = Remdata[IndexNew2,]
test_data2 <-Remdata [-IndexNew2,]
head(test_data2)
```
#Here is where I will perform normalization
```{r}
norm_values2 <-preProcess(training_data2,method= c ("center", "scale"))
train_nf2 <- predict(norm_values2, training_data2)
validation_nf2 <-predict(norm_values2, validation_data2)
test_nf2 <-predict(norm_values2,test_data2)
total_nf2 <-predict(norm_values2,UBank)
training_data2 <- train_nf2 [,-7]
training_outcome2 <- factor(validation_data2[,7],levels = C(0,1),labels = C( "Deny","Accept"))
valid_data2<-validation_nf2 [,-7]
validation_outcome2 <- factor(validation_data2[,7], levels= C(0,1), labels = C("Deny", "Accept"))
TestTraining_data2 <-test_nf2[,-7]
TestTraining_Outcome2 <-factor(TestTraining_data2[,7], levels= C(0,1), labels = C("Deny", "Accept")) 
Total_data2 <- total_nf2 [,7]
total_outcome2 <-factor(UBank[,7], levels = C(0,1), labels = C("Deny", "Accept"))
```
#Applying KNN with the optimal value of (K=4) to the Training and Validation Set #Validation
```{r}
knn_validation <-knn(train_data2,valid_data2,cl=training_outcome2,k=bestk_fit$k,prob= TRUE)
crossTable(validation_outcome2,knn_validation,prop.chisq = FALSE)

```
#Applying KNN with the optomal value of (K=4) to the Training and Test Set
#Test
```{r}
knn_testing <-Knn(training_data2,TestTraining_data2, cl= training_outcome2, k=bestk_fit$k, prob = TRUE)
CrossTable (TestTraining_Outcome2,knn_testing,prop.chisq= FALSE)
```
#Applying KNN with the optimnal value of (K=4) to the entire dataset 
#Total
```{r}
knn_total<- knn(train_data2, Total_data2, cl= training_outcome2, k=bestk_fit$k,prob=TRUE)
CrossTable(total_outcome2,knn_total,prop.chisq= FALSE)
```
# The difference between repartition data and the confusion matrix is that the data appears to be more accurate as the optimal value of k was applied. 